---
title: "vector_hw"
format: 
  html:
    toc: true
editor: source
---

## Vector Homework Notes:

  *Assign vectors using "c()" function
  
  *The function class() indicates what kind of object you are working with
  
  *The function str() provides an overview of the structure of an object and its elements


### Atomic Vector Types:
  1. Numeric
  2. Character
  3. Logical
  4. Integer
  5. Complex
  6. Raw
  
## Challenge 1:
  - What happens when we mix atomic vector types in a single vector?
```{r}
weight <- c(2,2L,1+4i)
```
* Created values for weight, 
  - all complex- converted all to be the same type of data

##Challenge (cont):
What happens in these examples?
```{r}
num_char <- c(1, 2, 3, "a") 
class(num_char)
```
  *Assigned character to this example- because that was the final piece of data(?)
  
  
```{r}
num_logical <- c(1, 2, 3, TRUE)
class(num_logical)
```
*Assigned numeric, is this a mistake? Why isn't it logical?


```{r}
char_logical <- c("a", "b", "c", TRUE)
```
*Assigned character, so it prefers num/ char to logical?


```{r}
tricky <- c(1, 2, 3, "4")
```
*Assigned character to "tricky," so it is "chr" when numeric value is in ""

  * *trying to find a common denominator, but is it just random?*
  

##Challenge (cont 2):
How many values are TRUE in combined data?
```{r}
combined_logical <- c(num_logical, char_logical)
```
*1 TRUE because it was deemed a character in char_logical, and replaced with a numeric value in num_logical

##Hierarchy drawing:
logical -> numeric -> character

###Subsetting Vectors:
```{r}
colors <- c("blue","green","red","yellow")
colors[3]

more_colors <- c("blue","green","red","yellow","purple","pink")
```

##Challenge (optional):
Why is "four" > "five" TRUE?
  * In string/ char form it is looking at alphabetical order- because four falls below five alphabetically it is considered "greater than"
